Analyzing the impact of social attributes on the success of commit integration

### 2017-02-18 Paper/MSRChallenge2017.tex
editted the results

\section{Building and Augmenting dataset with social attributes} 
\label{approach}
For this section our goal is to build a dataset based on the attributes that are 
intrinsic for the commit complexity and therefore for its probability of leading 
to a successful Travis CI job.
To do this we extracted the attributes from the original dataset 
provided~\cite{msr17challenge} which work as a proxy for the complexity of the 
commit being pushed.
\begin{itemize}
%\item tr\_original\_commit
%\item gh\_build\_started\_at 
\item git\_diff\_src\_churn 
\item gh\_diff\_files\_added 
\item gh\_diff\_files\_deleted 
\item gh\_diff\_files\_modified 
\item gh\_diff\_tests\_added 
\item gh\_diff\_tests\_deleted 
\item gh\_diff\_src\_files 
\item gh\_sloc 
%\item tr\_status 
\end{itemize}

Once we had these attributes, we removed repeated rows and we gathered the 
social attributes for each of these rows.

To do this, we extracted from the original file provided~\cite{msr17challenge} 
the commit hash and the project name. We removed the repeated rows and created a 
script to gather the social data of the developer who pushed that commit.
First we concatenated the two attributes before mentioned in the following way:
\newline \newline
\textit{https://github.com/\textbf{NameOfProject}/commit/\textbf{CommitHash}}
\newline \newline
We created a script that constructs this string and navigates through the html 
representation of the website described above, and looks for the 
\textit{username} of the developer in the Github platform. If it is able to find 
one, we then create another string following the pattern:
\newline \newline
\textit{https://github.com/\textbf{UserName}}
\newline \newline
We then navigate through the html representation of the developer's profile 
website (described above) and gather the following attributes:
\begin{itemize}
\item Repositories
\item Stars
\item Followers
\item Following
\item Contributions 
\end{itemize}

Repositories alludes to the repositories the developer has access to. This 
includes where the developer is the repository owner,where the developer is a 
collaborator, and repositories that the developer can access through an 
organization membership.\footnote{https://developer.github.com/v3/repos/} Stars 
show the projects the developer has stared, where starring translates to showing 
appreciation to the repository and creating a bookmark for later 
access.\footnote{https://help.github.com/articles/about-stars/} Followers and 
Following detail the number of developers that are following and being followed 
by the developer in question. Finally, the last attribute studied is the number 
of contributions made by the developer in the last year. To this dataset, we 
also added from the original dataset provided~\cite{msr17challenge} the 
attribute gh\_by\_core\_team\_member which we considered to have also social 
weight, which describes if the developer pushing the commit is a core member of 
the project the commit is being pushed to.

After performing this last data mining task, we created two different datasets. 
One of them contains technical data that is seen as a proxy for the complexity 
of the changes being performed in the commit. We also created a second dataset 
which consists of the data mentioned in the first dataset, plus the set of 
social attributes related to the developer who pushed the commit.

The dataset used in this study~\cite{msr17challenge} is very diverse 
and consists of commits that date from far back, and the social data we are 
gathering is recent data, gathered in January 2017. The researchers of this 
project acknowledge this discrepancy and in an effort to mitigate this threat to 
validity, we filter the commits analyzed by looking only at the commits pushed 
in the last 3 years of the dataset (2014,2015,2016), with the intuition that the 
social data gathered by this study will have a closer relationship to the social 
data if it were to be gathered in these years, than if we would compare it to 
years further back.

We constructed a decision tree with the dataset before mentioned, and we noticed that at this point the training data was unbalanced in a ratio close to 3:1 regarding successful and unsuccessful instances accordingly. To mitigate the risk of this being a skewing factor for the training process of the decision tree, we sampled randomly the successful instances to get an equal distribution of 
successful and unsuccessful instances to train the decision tree. We present the results of both of these datasets in the following section.

\section{Evaluation}
\todo{I guess we have to remove the logisitic regression section here if we are not going to present those results. Depending on what Bogdan and Claire think.}
\todo{consider moving this data preprocessing step to the previous section}
To explore how much social aspects contributed to the build success, we first
converted the build statuses to successful (passed) and not successful (errored,
failed, canceled, and started).  We then created a decision tree classifier
using the technical data to establish how influential the technical data is on 
build success.  We then added the social data to the technical data and used
that to create a new decision tree classifier.  The decision trees were
evaluated using ten fold cross validation.

\todo{this paragraph is now out of place.  Decide where to move it.}
We also evaluate our dataset by using decision trees, a predictive model which 
branches the different possible class values (represented as the leaves) 
depending on the attributes used to predict the class value (represented as the 
branches). In classification trees as the one being used in this study, each 
attribute represents a branch (with some exceptions due to pruning) and 
depending on the decisions taken while traversing the tree, the class value will 
be determined when one of the decisions leads to a leaf of the tree, therefore a 
class value will be encountered and a prediction can be made.

@@

Table~\ref{resultsTable} shows the results of running the data sets through the
default Weka J48 decision tree algorithm. 
In the original data set, 74.43\% of the builds were successful, the rest were not
successful.  This means that a classifier which guesses successful every time
would produce a classification rate of 74.43\%. The decision tree built from
only the technical indicators has a correct classification rate of 77.01\%, so
the technical attributes only increased the prediction accuracy by about 3\%.
Adding the social data increased the classifier's accuracy from 77.01\% to
81.10\%, about a 4\% increase.

@@

The same

@@

\section{Discussion}
In this study we have analyzed how considering social data is beneficial to be able to predict if a commit will be successfully integrated into a project or not. We have analyzed different technical complexity measures that work as a proxy for the complexity of the task being commited, such as number of files added, size of the project, size of the commit, etc. It is worth noticing that these are not the only measures that can be used as a proxy for technical complexity of a commit. The same intuition applies to the social attributes we gathered. These are not the only proxies for experience a developer may have, but we are making use of a suitable feature Github provides, which are the social attributes being measured in this study.

This study is consistent with previous studies~\cite{Yu15,tsay14icse,tsay14fse} that show the importance of considering social traits  when dealing with a merely social process such as the software development process.

@@

\section{Threats to validity}

\noindent\textbf{Internal validity:}
The internal validity threats deal with possible errors in our implementation 
and experiments.
The researchers in this study used the dataset provided by the 
MSRChallenge~\cite{msr17challenge} to gain useful information regarding the 
behavior of developers when pushing commits and how well these commits integrate 
successfully with the project. The authors have used 
Weka,\footnote{http://www.cs.waikato.ac.nz/ml/weka/} the well known tool for 
usage of machine learning algorithms to analyze our datasets. We are aware that, 
as every human written tool, Weka is prone to contain errors and/or bugs; 
nonetheless, Weka is a broadly used, open source tool that is subject to high 
quality standards and a vast user and developer base which makes it more likely 
to be a high quality code and to have less bugs than lesser used tools.

>>>>

\section{Building Dataset}
\label{approach}
To investigate our research questions, we built a dataset based on the 
attributes that consisted of both technical and social factors.  We collected
the technical factors from the provided data set~\cite{msr17challenge}.
We extracted only the attributes in the data set that proxy the commit's
complexity and the build status.
\begin{itemize}
%\item tr\_original\_commit
%\item gh\_build\_started\_at 
\item git\_diff\_src\_churn 
\item gh\_diff\_files\_added 
\item gh\_diff\_files\_deleted 
\item gh\_diff\_files\_modified 
\item gh\_diff\_tests\_added 
\item gh\_diff\_tests\_deleted 
\item gh\_diff\_src\_files 
\item gh\_sloc 
\item tr\_status 
\end{itemize}

Once we had these attributes, we removed repeated rows \todo{Why did we have to
remove repeated rows?}

To collect the social data, we created a web crawler to extract the social data
from developer's profiles.
We used the commit hash and the project name in the original data
set~\cite{msr17challenge} to locate the webpage of the commit using the syntax
below. 
\newline \newline
\textit{https://github.com/\textbf{NameOfProject}/commit/\textbf{CommitHash}}
\newline \newline
Then, we scraped the commit author's
\textit{username} from the page. If the scrapper was able to find username, the
scraper went to the user's profile information using the syntax below.
\newline \newline
\textit{https://github.com/\textbf{UserName}}
\newline \newline
The scrapper then extracted the following items from the user's profile.
\begin{itemize}
\item Repositories
\item Stars
\item Followers
\item Following
\item Contributions 
\end{itemize}

Repositories represents the repositories the developer has access to. This 
includes where the developer is the repository owner,where the developer is a 
collaborator, and repositories that the developer can access through an 
organization membership.\footnote{https://developer.github.com/v3/repos/} Stars 
represents a count of the projects the developer has stared. Starring  a project
means to showing approval of the repository and creating a bookmark for later 
access.\footnote{https://help.github.com/articles/about-stars/} Followers and 
Following detail the number of developers that are following and being followed 
by the developer in question. Finally, the last attribute studied is the number 
of contributions made by the developer in the last year.

We also added one item from the original data set to the social factors:
gh\_by\_core\_team\_member.
This factor indicates if the developer making a contribution is a core member 
of that project.

We used this data to create two data sets.  One with only the technical
factors and the build status, and another with both the technical and social
factors along with the build status.

The dataset used in this study~\cite{msr17challenge} is very diverse 
and consists of commits that date from far back, and the social data we are 
gathering is recent data, gathered in January 2017. The researchers of this 
project acknowledge this discrepancy and in an effort to mitigate the threat of
using future result to predict previous events,
we removed the commits in the dataset that were not performed within the last
three years of the data set's release (2014,2015,2016).  While this is not
perfect, we believe that these numbers are unlikely to have changed
significantly in the last three years.

After building the data set, we converted the various Travis build statuses to
successful (passed in the data set) or not successful (errored, failed, canceled
and started in the data set).  This still left a big discrepancy in the two
categories, with about a 3:1 successful to not successful ratio. To avoid having
this discrepancy alter our results, we created a separate data set where the
successful builds were randomly selected until the number of successful builds
equaled the number of not successful builds (downsampled).  We used both the
downsampled and non-downsampled data sets in our results.

\section{Evaluation}

We evaluated the predictive power of the social factors using a decision
tree.\todo{citation needed}
Decision trees are a predictive model which uses binary decisions on the
attributes to predict a final class.  The predicted class is represented as the
leaves, while the attribute decisions used in the prediction cause the
different branches.
In this study, we used the J48 Weka decision tree \todo{citation needed} with default settings. The 
decision trees were evaluated using ten fold cross validation with the data in
each data set.

@@

In the original data set, 74.43\% of the builds were successful, the rest were not
successful.  This means that a classifier which guesses successful every time
would produce a classification rate of 74.43\% if ran on the full data set.

Table~\ref{resultsTable} shows the results of running the two data sets through the
default Weka J48 decision tree algorithm.The decision tree built from
only the technical indicators has a correct classification rate of 77.01\% which
means the technical attributes only increased the prediction accuracy by about 3\%.
Adding the social data increased the classifier's accuracy from 77.01\% to
81.10\%, about a 4\% increase.

@@

@@

\section{Discussion}
We have found that social attributes which proxy experience produce a small but
noticeable increase in predicting build success.  While the increase was very
small with the full data set, there was a greater increase in prediction
accuracy after performing downsampling.  These results show that experience is
important to predicting build commit success but a strong predictor.

The build success by core team member results show that core team members are
more likely to have the build integration pass than non-core team members.  This
result provides evidence that non-core team member commit are not throughly
scrutinized, although further studied will be required to draw definite
conclusions.

@@

\section{Threats to validity}
This study contains various threats to validity.

\noindent\textbf{Internal validity:}
The researchers in this study used the dataset provided by the 
MSRChallenge~\cite{msr17challenge} to gain useful information regarding the 
behavior of developers when pushing commits and how well these commits integrate 
successfully with the project. The authors have used 
Weka,\footnote{http://www.cs.waikato.ac.nz/ml/weka/} the well known tool for 
usage of machine learning algorithms to analyze our datasets. We are aware that, 
as every human written tool, Weka is prone to contain errors and/or bugs; 
nonetheless, Weka is a broadly used, open source tool that is subject to high 
quality standards and a vast user and developer base which makes it more likely 
to be a high quality code and to have less bugs than lesser used tools.
